Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2025-03-26 paper.qmd
Add
Figure 18 updated
small typo

two_curvy_proj_tsne_first_model1 <- plot_proj(
    projected_df = projected_df, 
    model_df = model_df, 
    axes = axes, 
    circle = circle, 
    point_param = c(0.5, 0.2, "#636363"), # size, alpha, color
    line_param = c(0.5, 0.5, "#000000"), # linewidth, alpha
    plot_limits = c(-0.6, 0.6), 
    title = "c", 
    cex = 1.5, 
    position = sc_ltr_pos,
    axis_text_size = 4, 
    is_category = FALSE) +
    theme(
        legend.position = "none"
    )

two_curvy_proj_true <- plot_proj(
  projected_df = projected_df, 
  model_df = true_model_df_proj, 
  axes = axes, 
  circle = circle, 
  point_param = c(0.5, 0.5, "#636363"), # size, alpha, color
  line_param = c(0.5, 0.5, "#2b8cbe"), # linewidth, alpha
  plot_limits = c(-0.6, 0.6), 
  title = "a", 
  cex = 1.5, 
  axis_text_size = 4, 
  is_category = FALSE) +
  theme(
    legend.position = "none"
  )

@@

two_curvy_proj_true2 <- plot_proj(
    projected_df = projected_df, 
    model_df = true_model_df_proj, 
    axes = axes, 
    circle = circle, 
    point_param = c(0.5, 0.5, "#636363"), # size, alpha, color
    line_param = c(0.5, 0.5, "#2b8cbe"), # linewidth, alpha
    plot_limits = c(-0.42, 0.42), 
    title = "b", 
    cex = 1.5, 
    axis_text_size = 4, 
    is_category = FALSE) +
    theme(
        legend.position = "none"
    )

@@

trimesh_two_curvy_removed2_umap <- ggplot() +
  geom_segment(data = tr_from_to_df_two_curvy4,
               aes(
                 x = x_from,
                 y = y_from,
                 xend = x_to,
                 yend = y_to),
               colour = "#000000",
               alpha = 0.5,
               size = 0.5)+
  geom_point(data = umap_two_curvy_scaled,
             aes(
               x = UMAP1,
               y = UMAP2
             ),
             color = "#636363",
             alpha = 0.5,
             size = 0.5
  ) +
  coord_equal() +
  xlim(c(0, 1)) +
  ylim(c(0, 1)) +
  interior_annotation("b1", sc_ltr_pos, cex = 2)

@@

trimesh_two_curvy_removed1_umap <- ggplot() +
  geom_segment(data = tr_from_to_df_two_curvy3,
               aes(
                 x = x_from,
                 y = y_from,
                 xend = x_to,
                 yend = y_to),
               colour = "#000000",
               alpha = 0.5,
               size = 0.5)+
  geom_point(data = umap_two_curvy_scaled,
             aes(
               x = UMAP1,
               y = UMAP2
             ),
             color = "#636363",
             alpha = 0.5,
             size = 0.5
  ) +
  coord_equal() +
  xlim(c(0, 1)) +
  ylim(c(0, 1)) +
  interior_annotation("c1", sc_ltr_pos, cex = 2)
```

@@

two_curvy_proj_model_delaunay_umap <- plot_proj(
    projected_df = projected_df, 
    model_df = model_df1, 
    axes = axes, 
    circle = circle, 
    point_param = c(0.5, 0.2, "#636363"), # size, alpha, color
    line_param = c(0.5, 0.5, "#000000"), # linewidth, alpha
    plot_limits = c(-0.6, 0.6), 
    title = "a2", 
    cex = 2, 
    position = sc_ltr_pos,
    axis_text_size = 4, 
    is_category = FALSE) +
    theme(
        legend.position = "none"
    )

two_curvy_proj_model_benchmark2_umap <- plot_proj(
    projected_df = projected_df, 
    model_df = model_df2, 
    axes = axes, 
    circle = circle, 
    point_param = c(0.5, 0.2, "#636363"), # size, alpha, color
    line_param = c(0.5, 0.5, "#000000"), # linewidth, alpha
    plot_limits = c(-0.6, 0.6), 
    title = "b2", 
    cex = 2, 
    position = sc_ltr_pos,
    axis_text_size = 4, 
    is_category = FALSE) +
    theme(
        legend.position = "none"
    )

@@

error_plot_one_curvy_hist_selected <- ggplot(error_df_one_curvy_abs, aes(x = sqrt_row_wise_total_error, y = density,
                                                                         colour = factor(error_cat_n,
                                                                                         levels = c("high", "low")))) +
  geom_point() +
  scale_color_manual(values=c('#800026', '#d9d9d9')) +
  xlab(expression(e[hj])) +
  ylab("") +
  interior_annotation("b1",
                      position = c(0.95, 0.9),
                      cex = 1.5) +
  theme_bw() +
  theme(aspect.ratio = 1,
        axis.title = element_text(size = 18),
        axis.text = element_text(size = 15))

error_plot_one_curvy_hist <- ggplot(error_df_one_curvy_abs,
                                    aes(x = sqrt_row_wise_total_error, y = density,
                                        colour = factor(error_cat_n2,
                                                        levels = c("high", "low")))) +
  geom_point() +
  scale_color_manual(values=c('#800026', '#d9d9d9')) +
  xlab(expression(e[hj])) +
  ylab("") +
  interior_annotation("a1",
                      position = c(0.95, 0.9),
                      cex = 1.5) +
  theme_bw() +
  theme(aspect.ratio = 1,
        axis.title = element_text(size = 18),
        axis.text = element_text(size = 15))
```

```{r}
#| label: link-model-one-dens-clust
plot_tsne_dens <- error_df_one_curvy_abs |>
  ggplot(aes(x = tSNE1,
             y = tSNE2,
             colour = factor(error_cat_n2,
                             levels = c("high", "low")))) +
  geom_point(alpha=0.7) +
  scale_color_manual(values=c('#800026', '#d9d9d9')) +
  theme(aspect.ratio = 1) +
  interior_annotation("a2",
                      position = c(0.08, 0.9),
                      cex = 1.5)

plot_tsne_dens_selected <- error_df_one_curvy_abs |>
  ggplot(aes(x = tSNE1,
             y = tSNE2,
             colour = factor(error_cat_n,
                             levels = c("high", "low")))) +
  geom_point(alpha=0.7) +
  scale_color_manual(values=c('#800026', '#d9d9d9')) +
  theme(aspect.ratio = 1) +
  interior_annotation("b2",
                      position = c(0.08, 0.9),
                      cex = 1.5)

@@

## To add error category
projected_df <- projected_df |>
  mutate(cluster = error_df_one_curvy_abs$error_cat_n2)

five_c_shaped_proj_tsne_model1 <- plot_proj(
    projected_df = projected_df, 
    model_df = model_df, 
    axes = axes2, 
    circle = circle, 
    point_param = c(1.5, 0.5), # size, alpha, color
    line_param = c(0.8, 0.4, "#000000"), # linewidth, alpha
    plot_limits = c(-0.35, 0.35), 
    title = "a3", 
    cex = 1.5,
    position = c(0.08, 0.9),
    axis_text_size = 4, 
    is_category = TRUE) +
    scale_color_manual(values = c('#800026', '#d9d9d9')) +
    theme(aspect.ratio = 1,
          legend.position = "none")

>>>>

two_curvy_proj_tsne_first_model1 <- plot_proj(
    projected_df = projected_df, 
    model_df = model_df, 
    axes = axes, 
    circle = circle, 
    point_param = c(0.5, 0.2, "#636363"), # size, alpha, color
    line_param = c(0.5, 0.5, "#000000"), # linewidth, alpha
    plot_limits = c(-0.6, 0.6), 
    title = "c", 
    cex = 1.5, 
    position = sc_ltr_pos,
    axis_text_size = 4, 
    is_category = FALSE) +
    theme(aspect.ratio = 1,
          legend.position = "none")

two_curvy_proj_true <- plot_proj(
  projected_df = projected_df, 
  model_df = true_model_df_proj, 
  axes = axes, 
  circle = circle, 
  point_param = c(0.5, 0.5, "#636363"), # size, alpha, color
  line_param = c(0.5, 0.5, "#2b8cbe"), # linewidth, alpha
  plot_limits = c(-0.6, 0.6), 
  title = "a", 
  cex = 1.5, 
  axis_text_size = 4, 
  is_category = FALSE) +
  theme(aspect.ratio=1, 
        legend.position = "none")

@@

two_curvy_proj_true2 <- plot_proj(
    projected_df = projected_df, 
    model_df = true_model_df_proj, 
    axes = axes, 
    circle = circle, 
    point_param = c(0.5, 0.5, "#636363"), # size, alpha, color
    line_param = c(0.5, 0.5, "#2b8cbe"), # linewidth, alpha
    plot_limits = c(-0.42, 0.42), 
    title = "b", 
    cex = 1.5, 
    axis_text_size = 4, 
    is_category = FALSE) +
    theme(aspect.ratio = 1, 
          legend.position = "none")

@@

trimesh_two_curvy_removed2_umap <- ggplot() +
  geom_segment(data = tr_from_to_df_two_curvy4,
               aes(
                 x = x_from,
                 y = y_from,
                 xend = x_to,
                 yend = y_to),
               colour = "#000000",
               alpha = 0.5,
               linewidth = 0.5)+
  geom_point(data = umap_two_curvy_scaled,
             aes(
               x = UMAP1,
               y = UMAP2
             ),
             color = "#636363",
             alpha = 0.5,
             size = 0.5
  ) +
  coord_equal() +
  xlim(c(0, 1)) +
  ylim(c(0, 1)) +
  interior_annotation("b1", sc_ltr_pos, cex = 2)

@@

trimesh_two_curvy_removed1_umap <- ggplot() +
  geom_segment(data = tr_from_to_df_two_curvy3,
               aes(
                 x = x_from,
                 y = y_from,
                 xend = x_to,
                 yend = y_to),
               colour = "#000000",
               alpha = 0.5,
               linewidth = 0.5)+
  geom_point(data = umap_two_curvy_scaled,
             aes(
               x = UMAP1,
               y = UMAP2
             ),
             color = "#636363",
             alpha = 0.5,
             size = 0.5
  ) +
  coord_equal() +
  xlim(c(0, 1)) +
  ylim(c(0, 1)) +
  interior_annotation("c1", sc_ltr_pos, cex = 2)
```

@@

two_curvy_proj_model_delaunay_umap <- plot_proj(
    projected_df = projected_df, 
    model_df = model_df1, 
    axes = axes, 
    circle = circle, 
    point_param = c(0.5, 0.2, "#636363"), # size, alpha, color
    line_param = c(0.5, 0.5, "#000000"), # linewidth, alpha
    plot_limits = c(-0.6, 0.6), 
    title = "a2", 
    cex = 2, 
    position = sc_ltr_pos,
    axis_text_size = 4, 
    is_category = FALSE) +
    theme(aspect.ratio = 1, 
          legend.position = "none")

two_curvy_proj_model_benchmark2_umap <- plot_proj(
    projected_df = projected_df, 
    model_df = model_df2, 
    axes = axes, 
    circle = circle, 
    point_param = c(0.5, 0.2, "#636363"), # size, alpha, color
    line_param = c(0.5, 0.5, "#000000"), # linewidth, alpha
    plot_limits = c(-0.6, 0.6), 
    title = "b2", 
    cex = 2, 
    position = sc_ltr_pos,
    axis_text_size = 4, 
    is_category = FALSE) +
    theme(aspect.ratio = 1, 
          legend.position = "none")

@@

error_plot_one_curvy_hist_selected <- ggplot(error_df_one_curvy_abs, 
      aes(x = sqrt_row_wise_total_error, 
          y = density,
          colour = factor(error_cat_n, levels = c("low", "high")))) +
  geom_point(alpha=0.7) +
  scale_color_manual(values=c('#d9d9d9', '#800026')) +
  xlab(expression(e[hj])) +
  ylab("") +
  interior_annotation("b1",
                      position = c(0.95, 0.9),
                      cex = 1.5) +
  theme_bw() +
  theme(aspect.ratio = 1,
        axis.title = element_text(size = 18),
        axis.text = element_text(size = 15))

error_plot_one_curvy_hist <- ggplot(error_df_one_curvy_abs,
        aes(x = sqrt_row_wise_total_error, 
            y = density,
            colour = factor(error_cat_n2, levels = c("low", "high")))) +
  geom_point(alpha=0.7) +
  scale_color_manual(values=c('#d9d9d9', '#800026')) +
  xlab(expression(e[hj])) +
  ylab("") +
  interior_annotation("a1",
                      position = c(0.95, 0.9),
                      cex = 1.5) +
  theme_bw() +
  theme(aspect.ratio = 1,
        axis.title = element_text(size = 18),
        axis.text = element_text(size = 15))
```

```{r}
#| label: link-model-one-dens-clust
plot_tsne_dens <- error_df_one_curvy_abs |>
  ggplot(aes(x = tSNE1,
             y = tSNE2,
             colour = factor(error_cat_n2,
                             levels = c("low", "high")))) +
  geom_point(alpha=0.7) +
  scale_color_manual(values=c('#d9d9d9', '#800026')) +
  theme(aspect.ratio = 1) +
  interior_annotation("a2",
                      position = c(0.08, 0.9),
                      cex = 1.5)

plot_tsne_dens_selected <- error_df_one_curvy_abs |>
  ggplot(aes(x = tSNE1,
             y = tSNE2,
             colour = factor(error_cat_n,
                             levels = c("low", "high")))) +
  geom_point(alpha=0.7) +
  scale_color_manual(values=c('#d9d9d9', '#800026')) +
  theme(aspect.ratio = 1) +
  interior_annotation("b2",
                      position = c(0.08, 0.9),
                      cex = 1.5)

@@

## To add error category
projected_df <- projected_df |>
  mutate(cluster = factor(error_df_one_curvy_abs$error_cat_n2,
                          levels=c("low", "high")))

five_c_shaped_proj_tsne_model1 <- plot_proj(
    projected_df = projected_df, 
    model_df = model_df, 
    axes = axes2, 
    circle = circle, 
    point_param = c(1.5, 0.5), # size, alpha, color
    line_param = c(0.8, 0.4, "#000000"), # linewidth, alpha
    plot_limits = c(-0.35, 0.35), 
    title = "a3", 
    cex = 1.5,
    position = c(0.08, 0.9),
    axis_text_size = 4, 
    is_category = TRUE) +
    scale_color_manual(values = c('#d9d9d9', '#800026')) +
    theme(aspect.ratio = 1,
          legend.position = "none")
